Intrinsic Verification of Parsers and Formal Grammar Theory in Dependent Lambek Calculus

### 2024-11-14 paper/paper.tex
lambek

\section{Discussion and Future Work}
\label{sec:discussion}
\subsection{Typed Approaches to Grammars}

%% \subsubsection{Lambek Calculus and Categorial Grammar}
Lambek's original syntactic calculus \cite{lambek58} describes a logical system
for linguistic derivations. Later, he explicitly gives semantics to his
syntactic calculus inside of a non-commutative biclosed monoidal category
\cite{lambek1988categorial}. This led to many uses of non-commutative linear logic and lambda calculi in
linguistics \cite{buszkowskiTypeLogicsGrammar2003}. Grammars based on logic in
this manner may go by the names Lambek calculus, typelogical grammar, or
categorial grammar. Such categorial grammars correspond to the subtheory of our
calculus generated by $\otimes$, $\lto$, and $\tol$.

While applicable to both
formal and natural languages, the existing works on categorial grammar are
frequently oriented toward giving functorial semantics to natural language in
the tradition of Montague \cite{Montague1974-MONFPS}. The Abstract Categorial
Grammar Toolkit \cite{Guillaume2024} and Grammatical Framework \cite{ranta-2011}
each mechanize aspects of Lambek calculus ; however, to the best of our knowledge,
\theoryabbv is the first mechanization designed for formal language parsing first.
Moreover, we provide a novel use of inductive and dependent types inside of a
Lambek calculus.

>>>>

\section{Discussion and Future Work}
\label{sec:discussion}
\subsection{Typed Approaches to Grammars}
\subsubsection{Lambek Calculus and Categorial Grammar}
Lambek's original syntactic calculus \cite{lambek58} describes a logical system
for linguistic derivations, and it can be given semantics inside of a non-commutative biclosed monoidal category
\cite{lambek1988categorial}. This led to many uses of non-commutative linear logic and lambda calculi in
linguistics \cite{buszkowskiTypeLogicsGrammar2003} --- including mechanized
categorial grammar parsers \cite{Guillaume2024,ranta-2011}. This style of grammar
formalism has gone by the names Lambek calculus or \emph{categorial grammar},
and it is equal in expressivity to context-free grammars.
The existing works on categorial grammar
frequently present logical operations on linear predicates rather than the term
language given in this paper.
Additionally, much of the existing work is
oriented toward computational linguistics, giving functorial semantics to natural language in
the tradition of Montague \cite{Montague1974-MONFPS}.
